---
title: "Wolt Customer Segmentation Analysis"
author: "Alexander Ayupov"
date: "`r Sys.Date()`"
output:  
  html_document: default
  html_notebook: default
---


# Wolt Customer Segmentation Analysis
For analysis, the method of customer segmentation by clusters will be used.
The reason for having this approach is that dataset contains customer database, and to make sense of data it's reasonable to breakdowm customers by set of similarities.
Analysis will result in having a defined customers groups to target.

## Load required packages
```{r load packages, echo=TRUE, results='hide', warning=FALSE, message=FALSE}
install.packages("tidyverse", repos = "http://cran.us.r-project.org")
install.packages("skimr", repos = "http://cran.us.r-project.org")
install.packages("janitor", repos = "http://cran.us.r-project.org")
install.packages("DT", repos = "http://cran.us.r-project.org")
install.packages("rmarkdown", repos ="http://cran.us.r-project.org")
install.packages("knitr", repos = "http://cran.us.r-project.org")
install.packages("plyr", repos = "http://cran.us.r-project.org")
library(DT)
library(tidyverse)
library(readr)
library(skimr)
library(janitor)
library(dplyr)
library(ggplot2)
library(tidyr)
library(knitr)
library(rmarkdown)
library(lubridate)
library(rmarkdown)
library(knitr)
```

## Import Data
```{r, echo=TRUE, results='hide', warning=FALSE, message=FALSE}
wolt_dataset <- read_csv("dataset_for_analyst_assignment_20201120.csv")
```

## Understanding dataset, preparations
```{r Get to know dataset, echo=TRUE, results='hide', warning=FALSE, message=FALSE}
glimpse(wolt_dataset)
as_tibble(wolt_dataset) %>% 
  print(n=50)

colnames(wolt_dataset)
summary(wolt_dataset)
```

```{r, echo=TRUE, results='hide', warning=FALSE, message=FALSE}
any(is.na(wolt_dataset))      #Check if there are null values
nrow(na.omit(wolt_dataset))
nrow(wolt_dataset)            #Check total number of observations
```

```{r, echo=TRUE, results='hide', warning=FALSE, message=FALSE}
sum(is.na(wolt_dataset$USER_ID))        #Check for null values in a "key" columns
sum(is.na(wolt_dataset$PURCHASE_COUNT))
length(unique(wolt_dataset$USER_ID)) == nrow(wolt_dataset) #Check for duplicates 
```

```{r, echo=TRUE, results='hide', warning=FALSE, message=FALSE}
length(unique(wolt_dataset$REGISTRATION_COUNTRY)) #Check unique values
```

## Data Cleaning

```{r, echo=TRUE, results='hide', warning=FALSE, message=FALSE}
wolt_dataset_cleaned <- wolt_dataset %>% filter(PURCHASE_COUNT>0)   #handling nulls
```

```{r, echo=TRUE, results='hide', warning=FALSE, message=FALSE}
wolt_dataset_cleaned <- wolt_dataset_cleaned %>%       
  filter(REGISTRATION_COUNTRY=="FIN" |
           REGISTRATION_COUNTRY=="DNK" |
           REGISTRATION_COUNTRY=="GRC")
```


## Group Customers by Registration country

```{r}
viz_by_country <- wolt_dataset_cleaned %>%
  group_by(REGISTRATION_COUNTRY) %>%                             
  summarize(customers = n()) %>%
  arrange(customers) %>%                                
  mutate(REGISTRATION_COUNTRY = factor(REGISTRATION_COUNTRY, REGISTRATION_COUNTRY)) %>%     
  ggplot(aes(x=REGISTRATION_COUNTRY, y=customers, fill=REGISTRATION_COUNTRY)) +                  
  geom_bar(stat="identity") +                      
  coord_flip() +
  theme(legend.position="none") +
  geom_text(aes(label = round(customers))) +
  labs(title ="Number of Customers/Country", caption = "By Alexander Ayupov", x="Country", y="Customers")
       
```


```{r }
plot(viz_by_country)
```

## Breakdown countries by Total purchases

```{r}
viz_country_vs_income <- wolt_dataset_cleaned %>%
  group_by(REGISTRATION_COUNTRY) %>% 
  summarize(pcount = sum(TOTAL_PURCHASES_EUR)) %>% 
  ggplot(aes(x = REGISTRATION_COUNTRY, y = pcount, fill = REGISTRATION_COUNTRY)) +
  geom_col() + 
  coord_flip() +
  theme(legend.position="none") +
  geom_text(aes(label = round(pcount))) +
  scale_y_continuous(labels = scales::dollar) +
  labs(title ="Total purchases by Country", caption = "By Alexander Ayupov", x="Country", y="Total Purchases")
```

```{r}
plot(viz_country_vs_income)
```

## Customers preferred device by country

```{r}
viz_prefdevice_by_country <- wolt_dataset_cleaned %>% 
  filter(!is.na(PREFERRED_DEVICE)) %>% 
  ggplot(aes(PREFERRED_DEVICE, fill=PREFERRED_DEVICE)) +
  geom_bar() +
  facet_wrap(~REGISTRATION_COUNTRY) +
  theme(legend.position="none") +
  labs(title ="Preferred Device by Country", caption = "By Alexander Ayupov", x="Device", y="Customers")
```

```{r}
plot(viz_prefdevice_by_country)
```

# Clustering Customers

## User Recency data transformation
```{r}
user_recency <- wolt_dataset_cleaned %>% 
  select(USER_ID, MEDIAN_DAYS_BETWEEN_PURCHASES) %>% 
  rename(recency = MEDIAN_DAYS_BETWEEN_PURCHASES) %>% 
  group_by(USER_ID)

user_recency[is.na(user_recency)] <- 0   #replace na with 0

user_recency_filter <- user_recency %>%     #filter outliers
  filter(recency <= 100)

print(summary(user_recency$recency)) 
```

## Plotting user recency, group customers by Inactivity period

```{r}
viz_recency <- ggplot(user_recency_filter, aes(recency)) +
  geom_histogram(bins = 20, color=4, fill="gray") +
  ylab('Number of Customers') +
  theme_minimal() +
  labs(title ="Day of Customers Inactivity", subtitle = "Order recency", caption = "By Alexander Ayupov", x="Days of Inactivity", y="Customers") +
  annotate("text", x=50,y=4000,label="50% of Users have ~9 days of Inactivity\nAverage is one month without making a single purchase", color="blue")
```

```{r}
plot(viz_recency)
```

## User Frequency data transformation

```{r}
user_frequency <- wolt_dataset_cleaned %>% 
  select(USER_ID, PURCHASE_COUNT) %>% 
  rename(Frequency = PURCHASE_COUNT) %>% 
  group_by(USER_ID)

user_frequency_filter <- user_frequency %>%     #filter outliers
  filter(Frequency <= 30)

print(summary(user_frequency_filter$Frequency))
```

## Plotting user frequency, group by transaction per customers

```{r}
viz_frequency <- ggplot(user_frequency_filter, aes(USER_ID, Frequency)) +
  geom_boxplot(color=4) +
  theme(axis.ticks.x = element_blank()) +
  theme_minimal() +
  labs(title ="Purchase Frequency", caption = "By Alexander Ayupov", x="Customers", y="Number of Purchases per Customer") +
  annotate("text", size=3.3, x=16700,y=25,label="Average 6 transaction per Customer\n50% of customers have ~3 transactions", color="blue")
```

```{r warning=FALSE}
plot(viz_frequency)
```

## User Monetary value data transformation
### Users summary less 1000€ value
```{r}
user_monetary_value <- wolt_dataset_cleaned %>% 
  select(USER_ID, TOTAL_PURCHASES_EUR) %>% 
  rename(monetary_value = TOTAL_PURCHASES_EUR) %>% 
  group_by(USER_ID)

MV_filter_1000 <- user_monetary_value %>% 
  filter(monetary_value <= 1000)

print(summary(MV_filter_1000$monetary_value))
```
### Users summary over 1000€ value
```{r}
MV_filter_over1000 <- user_monetary_value %>% 
  filter(monetary_value > 1000)

print(summary(MV_filter_over1000$monetary_value))
```

## Plotting user Monetary value
### Users monetary value less 1000€

```{r}
viz_monetary_1000 <- ggplot(MV_filter_1000, aes(monetary_value)) +
  geom_histogram(bins=10, color=4,fill="grey") +
  scale_x_continuous(labels = scales::dollar) +
  scale_y_continuous(labels = scales::comma) +
  labs(title ="Revenue of Users - Below €1K", caption = "By Alexander Ayupov", x="Revenue", y="Number of Customers") +
  annotate("text", x=650,y=4000,label="50% of Users brings 70€ revenue\nAverage revenue per user is 150€", color="blue")
```

```{r}
plot(viz_monetary_1000)
```

### Users monetary value over 1000€

```{r}
viz_monetary_ov1000 <- ggplot(MV_filter_over1000, aes(monetary_value)) +
  geom_histogram(bins=10, color=4,fill="grey") +
  scale_x_continuous(labels = scales::dollar) +
  scale_y_continuous(labels = scales::comma) + 
  labs(title ="Revenue of Users - Over €1K", caption = "By Alexander Ayupov", x="Revenue", y="Number of Customers") +
  annotate("text", x=6000,y=100,label="50% of Users brings 1350€ revenue\nAverage revenue per user is 1600€", color="blue")
```

```{r}
plot(viz_monetary_ov1000)
```

## Merging recency, frequence, monetary value subsets

```{r}
Users_RFM <- merge(user_recency, user_frequency) 
Users_RFM <- merge(Users_RFM, user_monetary_value) 
DT::datatable((Users_RFM),
              rownames = FALSE,
              options = list(
                pageLength = 10))
```

## Clustering Customers segments

```{r}
set.seed(415)
clusters <- kmeans(scale(Users_RFM[,2:4]), 3, nstart = 1)
Users_RFM$Cluster <- as.factor(clusters$cluster)

KMeans_Results <- Users_RFM %>%
  group_by(Cluster) %>%
  summarise('Number of Users' = n(),
            'Recency Mean' = round(mean(recency)),
            'Frequency Mean' = scales::comma(round(mean(Frequency))),
            'Monetary Value Mean' = scales::dollar(round(mean(monetary_value))),
            'Cluster Revenue' = scales::dollar(sum(monetary_value))
  )

table_clusters <- DT::datatable((KMeans_Results),
              rownames = FALSE)

DT::datatable((KMeans_Results),
              rownames = FALSE)
```


## Plotting Customers Clusters

```{r}
Cluster_size_visz <- ggplot(KMeans_Results, aes(Cluster, `Number of Users`)) +
  geom_text(aes(label = `Number of Users`), vjust = -0.3) +
  geom_bar(aes(fill=Cluster), stat='identity') +
  theme_classic() +
  labs(title ="Number of Users per Cluster", caption = "By Alexander Ayupov", x="Clusters", y="Number of Customers")
```

```{r}
plot(Cluster_size_visz)
```

# Breakdown Users data by Daytime/Weekdays

## Data transformation- Parsing Purchase hours 

```{r}
wolt_dataset_cleaned <- wolt_dataset_cleaned %>%
  mutate(MOST_COMMON_HOUR_OF_THE_DAY_TO_PURCHASE = parse_date_time(MOST_COMMON_HOUR_OF_THE_DAY_TO_PURCHASE, 'H:'))   #format as date
```

```{r}
wolt_dataset_cleaned <- wolt_dataset_cleaned %>%
  mutate(MOST_COMMON_HOUR_OF_THE_DAY_TO_PURCHASE = format(strptime(MOST_COMMON_HOUR_OF_THE_DAY_TO_PURCHASE, "%Y-%m-%d %H:%M:%S"),'%H:%M'))   #Leave H:M format
```

## Data transformation- Formatting Purchase weekdays

```{r, echo=TRUE, results='hide', warning=FALSE, message=FALSE}
unique(wolt_dataset_cleaned$MOST_COMMON_WEEKDAY_TO_PURCHASE)

wolt_dataset_cleaned$MOST_COMMON_WEEKDAY <- factor(wolt_dataset_cleaned$MOST_COMMON_WEEKDAY_TO_PURCHASE, levels=1:7,
                                                   labels=c("Monday", "Tuesday", "Wednesday",
                                                            "Thursday", "Friday", "Saturday", "Sunday"))
```

## Plotting Most common hours purchase by country
### FIN

```{r}
viz_hours_fin <- wolt_dataset_cleaned %>%
  filter(REGISTRATION_COUNTRY=="FIN") %>%  
  group_by(MOST_COMMON_HOUR_OF_THE_DAY_TO_PURCHASE) %>% 
  summarize(pcount = sum(PURCHASE_COUNT)) %>% 
  ggplot(aes(x=MOST_COMMON_HOUR_OF_THE_DAY_TO_PURCHASE, y = pcount, fill=pcount)) +
  geom_col() +
  theme(axis.text.x = element_text(angle = 45)) +
  theme(legend.position="none") +
  labs(title ="Number of orders by day time FIN", caption = "By Alexander Ayupov", x="Hours of Purchase", y="Purchase count")
```

```{r}
plot(viz_hours_fin)
```

### DNK

```{r}
viz_hours_dnk <- wolt_dataset_cleaned %>%
  filter(REGISTRATION_COUNTRY=="DNK") %>%  
  group_by(MOST_COMMON_HOUR_OF_THE_DAY_TO_PURCHASE) %>% 
  summarize(pcount = sum(PURCHASE_COUNT)) %>% 
  ggplot(aes(x=MOST_COMMON_HOUR_OF_THE_DAY_TO_PURCHASE, y = pcount, fill=pcount)) +
  geom_col() +
  theme(axis.text.x = element_text(angle = 45)) +
  theme(legend.position="none") +
  labs(title ="Number of orders by day time DNK", caption = "By Alexander Ayupov", x="Hours of Purchase", y="Purchase count")
```

```{r}
plot(viz_hours_dnk)
```

### GRC

```{r}
viz_hours_grc <- wolt_dataset_cleaned %>%
  filter(REGISTRATION_COUNTRY=="GRC") %>%  
  group_by(MOST_COMMON_HOUR_OF_THE_DAY_TO_PURCHASE) %>% 
  summarize(pcount = sum(PURCHASE_COUNT)) %>% 
  ggplot(aes(x=MOST_COMMON_HOUR_OF_THE_DAY_TO_PURCHASE, y = pcount, fill=pcount)) +
  geom_col() +
  theme(axis.text.x = element_text(angle = 45)) +
  theme(legend.position="none") +
  labs(title ="Number of orders by day time GRC", caption = "By Alexander Ayupov", x="Hours of Purchase", y="Purchase count")
```

```{r}
plot(viz_hours_grc)
```

## Plotting Most common weekdays purchase by country
### FIN

```{r}
viz_days_fin <- wolt_dataset_cleaned %>%
  filter(REGISTRATION_COUNTRY=="FIN") %>% 
  group_by(MOST_COMMON_WEEKDAY) %>% 
  summarize(pcount = sum(PURCHASE_COUNT)) %>% 
  ggplot(aes(x = MOST_COMMON_WEEKDAY, y = pcount, fill=pcount)) +
  geom_col() +
  labs(title ="Number of orders by Weekdays FIN", caption = "By Alexander Ayupov", x="Days of Purchase", y="Purchase count")
```

```{r}
plot(viz_days_fin)
```

### DNK

```{r}
viz_days_dnk <- wolt_dataset_cleaned %>%
  filter(REGISTRATION_COUNTRY=="DNK") %>% 
  group_by(MOST_COMMON_WEEKDAY) %>% 
  summarize(pcount = sum(PURCHASE_COUNT)) %>% 
  ggplot(aes(x = MOST_COMMON_WEEKDAY, y = pcount, fill=pcount)) +
  geom_col() +
  labs(title ="Number of orders by Weekdays DNK", caption = "By Alexander Ayupov", x="Days of Purchase", y="Purchase count")
```

```{r}
plot(viz_days_dnk)
```

### GRC

```{r}
viz_days_grc <- wolt_dataset_cleaned %>%
  filter(REGISTRATION_COUNTRY=="GRC") %>% 
  group_by(MOST_COMMON_WEEKDAY) %>% 
  summarize(pcount = sum(PURCHASE_COUNT)) %>% 
  ggplot(aes(x = MOST_COMMON_WEEKDAY, y = pcount, fill=pcount)) +
  geom_col() +
  labs(title ="Number of orders by Weekdays GRC", caption = "By Alexander Ayupov", x="Days of Purchase", y="Purchase count")

```

```{r}
plot(viz_days_grc)
```